%-----------------------------------------------%
%             filename: skeleton.tex
%-----------------------------------------------%
\documentclass[aps,twocolumn]{revtex4-1}
\usepackage{graphicx}
\usepackage{ifpdf}
\ifpdf
	\usepackage[backref]{hyperref}
	%\usepackage[backref,pageanchor=true,plainpages=false, pdfpagelabels,bookmarks,bookmarksnumbered]	{hyperref}
\else
\fi

\usepackage{crs}

% see http://goo.gl/5Fo27
\newtoggle{thmsty}
%\toggletrue{thmsty}
\togglefalse{thmsty}

% see http://goo.gl/7jLZ9
\makeatletter
\newlength \figwidth
\if@twocolumn
  \setlength \figwidth {0.8\columnwidth}
\else
  \setlength \figwidth {0.5\textwidth}
\fi
\makeatother

\begin{document} 

\title{\bf Functional closure in biomolecular systems}

\author{a1$^{1}$, a2$^{1}$, a3$^{1}$, a4$^{1,2,3}$}

\affiliation{$^1$Department of Systems and Computational Biology,\\ $^2$Dominick P. Purpura Department of Neuroscience, \\ $^3$Department of Pathology, Albert Einstein College of Medicine, 1301 Morris Park Ave, Bronx, NY 10461, USA}

\date{\today}
\begin{abstract}
Hierarchy is a fundamental organizing concept in biology. The existence of a number of apparently discrete levels of organization including molecules, cells, organisms, populations, communities, ecosystems, and more fine-grained levels below, above and in between these are taken for granted. Theoretical constructs capable of integrating information about such levels of organization and of explaining the emergence of such levels in the evolutionary process have not been adapted to a biological context. Here we make use of existing theory to characterize coherence constraints that must ostensibly be satisfied in any evolutionary process in which two levels of organization become coupled such that one serves as a base supporting interactions among higher-level entities in another. In this way, what may be considered as two distinct levels may come to be viewed as part of a unified system. This process of unification of distinct levels may serve as a basis for induction in the consideration of arbitrarily complex compositions of multilevel biological systems. The language we use to express these constraints is sufficiently general that it applies equally to several different concrete formalisms for representing biological systems, for example those that have become commonplace in the context of network science, in terms of algebraic graphs, hypergraphs, or (cell) complexes. The derived relationship constraining the synthesis of levels of organization in models of biological systems can be used to provide boundary conditions for dynamical models defined in terms of transformations of graphs or any other concrete relational structure that seek to take into account hierarchical transitions in evolutionary processes.
\end{abstract}

\maketitle

\tableofcontents

\section{Introduction}

The current amount of biological data surpasses our present analytic and integrative capacity. The naturalist dream placing high value upon an extensive catalog of biological diversity has reached molecular resolution and yet we lack a fundamental theoretical framework in which to understand and interpret the relationships living systems exhibit that transcends levels of the biological hierarchy. Moreover, there is no proper language tailored to apprehend, communicate or abstractly settle an ensemble of biological problems that share a common structure. As the current paradigm apparently organized around data gathering continues, its limitations have become increasingly apparent. If we keep on characterizing our biosphere at the molecular level without a promising theoretical framework, we will soon become lost, if we are not already, in its vastness. What is worse, we could end up in biological Babel with no integrative perspective and thus no means of communication. Hence, from a purely pragmatic perspective, the scientific community should invest in the project of unification of knowledge and make the field of theoretical biology a socio-economically palatable career option for young and old scientists alike. In accordance with this thesis, the aim of this review is to consider a limited subset of past and present attempts to develop a formal language for biological systems focusing on the virtues and flaws of each approach in order to provide a reference for the current biologist interested in attempting to address the problem of understanding as it applies to biology.

Category theory has provided  a solid language in which to talk about mathematical systems and has been recently  used as such and such. In the beginning, category theory was considered by other mathematicians commonly as abstract nonsense but nowadays the community is starting to see it even in an applied fashion. Nevertheless, what cannot be criticized of this theoretical framework is its convenience as a formal language in order to deal with a collection of abstract systems such as mathematical theories. The fundamental organizational power of category theory resides in giving relevance to relationships over objects, thus obtaining an abstract framework into which unify different mathematical programs.

\section{The definition of living systems}

Rosen's metabolism and repair systems can be explained from a categorical perspective. In Life, Life itself, Rosen developed the concept of closure to efficient causation recovering partially the Aristotelian view of a causal explanation and by this giving some sense of structural definition of life. By differentiating between material and efficient cause, Rosen kept the nature of object and refined the concept of autopoiesis by Maturana and Varela. Invoking this characteristical clousure versus a material closure allows us to perfectly solve the parasite common issue.

\section{A scent of category theory}
It is not the objective of this work to give a detailed and more formal exposition of category theory but to provide the necessary perspective in order to understand the present text. As a reference in category theory the reader is advised to consult




%------bibliography---%
\bibliographystyle{unsrt} 
\bibliography{bib/books,bib/papers}
%---------------------%


%-----------------------------------------------%
%                   appendix
%-----------------------------------------------%
\appendix

\section{Category theory}\label{app:CatTh}
\input{tex/catth.tex}

%\section{Sieves and sheaves}\label{Sheaves}
%\input{sec/sievesheaf.tex}

\end{document}
